---
title: "Midterm presentation stat 641"
Group: 4
output: pdf_document
---


# Example 1:
# Description about the mouse dataset:
A small randomized experiment were done with 16 mice. 7 to treatment group and 9 to control group. Treatment was intended to prolong survival after a test surgery.
```{r}
library(bootstrap)
library(tidyverse)
library(infer)

# Data
treat <- mouse.t
ctrl <-  mouse.c

t1 <- length(treat)
t2 <- length(ctrl)

Group <- as.factor( c(rep(1, t1), rep(2, t2)))

Group <- as.factor(ifelse(Group == 1, "treat", "control"))


mouse <- data.frame(Group, Survival_time = c(treat, ctrl))

# Create an another variable called mouse1 which shuffles the original dataset.
mouse1 <- mouse %>% sample_n(size = 16)
head(mouse1,5)

# Exploratory data analysis
mouse1 %>%
  group_by(Group) %>%
  summarise(
    count = n(),
    Mean = round(mean(Survival_time),2)
  )
```
```{r}
# Visualize barplot using a ggplot2 package
mouse1 %>%
  ggplot(aes(x = Group, fill = Group)) + geom_bar() + labs(title = "Treatment Versus Control Group", x = "Groups", y = "Count") + scale_y_continuous(breaks = seq(from = 0, to = 10,by= 2))+theme(panel.grid.major = element_blank())

# Visualize 
mouse1 %>%
  ggplot(aes(x = Group,y = Survival_time)) + geom_point(aes(col = Group)) + labs(title = "Treatment Versus Control Group", x = "Groups", y = "Survival time (in days)") + scale_y_continuous(breaks = seq(from = 0, to = 200,by= 20)) +theme(panel.grid.major = element_blank())
```


```{r}
# Calculate an observed test statistic.
obs_test_stat <- mouse1 %>%
  specify(Survival_time ~ Group)%>%
  calculate(stat = "diff in means", order = c("treat","control"))
round(obs_test_stat,2)

# To double check manually computed observed test statistics.
obs_diff_mean = 86.86	- 56.22
obs_diff_mean
```

# Population of parameter interest: 
Difference in the population means 
$(\mu_{mt} - \mu_{mc})$

# Hypothesis testing:
(One-sided alternative)

* In words:

$H_0:$ There is no difference in the means of both groups.

vs.

$H_1:$ There is a significant difference in the means where the mice in the treatment group are favored.

* In notations:

$H_0 : \mu_{mt} - \mu_{mc} = 0$

vs.

$H_1 : \mu_{mt} - \mu_{mc} > 0$

# Point estimate/test statistics: 

* The point estimate is the difference in the sample means

$\hat{\bar x_{mt}} - \hat{\bar x_{mc}}$

or

${\hat\mu_{mt}} - {\hat\mu_{mc}}$

# Observed test statistics:
Observed difference in the means of both groups = 30.64.

# Null distribution
Null distribution is the sampling distribution of the test statistic assuming the null hypothesis is true. 

```{r}
# Create the null distribution.
null_dist <- mouse1 %>%
  specify(Survival_time ~ Group)%>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("treat","control"))
head(null_dist,5)

# Visualize the null distribution.
null_dist %>%
  visualize() +
  shade_p_value(obs_stat = obs_test_stat, direction = "right", col = "blue", lty = 1, lwd = 1)
```

# p-value: 
p-value is the probability of obtaining a test statistic just as extreme or more extreme than the observed test statistic assuming the null hypothesis is true. 
```{r}
# Compute the p-value.
null_dist %>%
  get_p_value(obs_stat = obs_test_stat, direction = "right")
```

Since the p-value > 0.05 (significance $\alpha$ level) we failed to reject the null hypothesis and concluded that there is no significant difference in the means of the both groups.

# EXAMPLE 2
A study in 1994 examined 491 dogs that had developed cancer and 945 dogs as a control group to determine whether there is an increased risk of cancer in dogs that are exposed to the herbicide 2,4-Dichlorophenoxyacetic acid (2,4-D).
```{r}
library(openintro)
cancer_in_dogs <- openintro::cancer_in_dogs
table(cancer_in_dogs)
dim(cancer_in_dogs)
```

```{r}
cancer_in_dogs$response <- ifelse(cancer_in_dogs$response == "no cancer", 0, 1)
```
```{r}
head(cancer_in_dogs,5)
```

```{r}
cancer_in_dogs %>%
  group_by(order) %>%
  summarise(
    count = n(),
    Mean = round(mean(response),2)
  )
# visualize
cancer_in_dogs %>%
  ggplot(aes(x = order,fill = response)) +geom_bar() + labs(fill = "Cancer type")

```

```{r}
# Calculate an observed test statistic.
obs_test_stat <- cancer_in_dogs %>%
  specify(response ~ order)%>%
  calculate(stat = "diff in means", order = c("2,4-D","no 2,4-D"))
round(obs_test_stat,2)

# To double check manually computed observed test statistics.
obs_diff_mean = 0.39	- 0.32
obs_diff_mean
```

# Population of parameter interest: 
Difference in the dog population means 
$(\mu_{1} - \mu_{2})$
$\mu_1 = dog \ population \ exposed \ to \ 2,4-D$
$\mu_2 = dog \ population \ exposed \ to \ no \ 2,4-D$

# Hypothesis testing:
(One-sided alternative)

In words:
$H_0:$ There is no difference in the mean of the dog population that are exposed to the herbicide 2,4,D and no 2,4,D.

vs.

$H_1:$ There is a significant difference in the means, where the dogs that are exposed to 2,4-D are favored.

In notations:
$H_0 : \mu_{1} - \mu_{2} = 0$

vs.

$H_1 : \mu_{1} - \mu_{2} > 0$

# Point estimate/test statistics: 
The point estimate is the difference in the sample means
$\hat{\bar{x}_{1}} - \hat{\bar{x}_{2}}$

or

${\hat\mu_{1}} - {\hat\mu_{2}}$

# Observed test statistics:
Observed difference in the means of both groups = 0.07.

# Null distribution
Null distribution is the sampling distribution of the test statistic assuming the null hypothesis is true. 

```{r}
# Create the null distribution.
null_dist <- cancer_in_dogs %>%
  specify(response ~ order)%>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("2,4-D","no 2,4-D"))
head(null_dist,5)

# Visualize the null distribution.
null_dist %>%
  visualize() +
  shade_p_value(obs_stat = obs_test_stat, direction = "right", col = "blue", lty = 1, lwd = 1)
```



```{r}
# Compute the p-value.
null_dist %>%
  get_p_value(obs_stat = obs_test_stat, direction = "right")
```

Since the p-value < 0.05 (significance $\alpha$ level) we reject the null hypothesis and support the alternative. We concluded that there is a significant difference in the means, where the dogs that are exposed to 2,4-D are favored.
























































